---
title: "ggplot Customization with National Park Visitation Data (Exercise)"
date: "2024-02-26"
categories: [ggplot, advanced, exercise]
format:
  html: default
code-overflow: wrap
code-fold: show
editor: visual
df-print: kable
R.options:
  warn: false
code-tools: true
execute:
  eval: false
---

# <span style="color:green;"> Exercises </span>

## ggplot Customization with National Park Visitation Data

<span style="color:red;"> [Solutions](Ggplot-Customization-NP-Solutions.qmd) </span>

These exercises use National Park visitation data from 1979–2024. For more context about the dataset, see the [data essay](../index.qmd).

**Concepts covered:**

- Filtering data for a specific category
- Line plots with custom colors and titles
- Customizing x-axis tick intervals
- Abbreviating y-axis labels (millions, thousands)
- Adjusting axis limits to zoom into a time period

## Load National Park Visitation data
```{r}
np_data <- read.csv("https://raw.githubusercontent.com/melaniewalsh/responsible-datasets-in-context/main/datasets/national-parks/US-National-Parks_RecreationVisits_1979-2024.csv",
 stringsAsFactors = FALSE)
```

View the np_data dataframe by clicking on the spreadsheet icon in the Global Environment

## Load libraries

```{r}
#| message: false
library("dplyr")
library("stringr")
library("ggplot2")
library("scales")
```

- How have visits to a particular National Park changed over time?
- What is the most interesting period of change?

# Exercise 1
First, filter the dataframe for a park of your choice. Then, pick a National Park that you haven't worked with yet, and filter the data for only that park.

```{r}
# Your code here
```

# Exercise 2

Now, make a line plot that shows the number of visits per year to that park from 1979 to 2022.

### 2a.
Choose a color for the line.

### 2b.
Give the plot a title that also functions as a kind of "headline" for the most interesting story of the plot.

### 2c.
Change the x-axis ticks so that they increase 5 years at a time.

### 2d.
Change the y-axis tick labels so that they abbreviate millions to M and thousands to K.


```{r}
# Your code here
```

## Exercise 3

Now, create a plot that zooms in on the most interesting time period for this particular National Park.

### 3a.
Change the x-axis limits so that it only shows the most interesting years.

### 3b.
Come up with a new title that describes this time period.


```{r}
# Your code here
```
