---
title: "DPLYR Groupby with National Park Visitation Data (Exercise)"
date: "2024-08-01"
categories: [dplyr, exercise]
format: 
  html: 
    code-links:
      - text: R Script
        href: NP-Data-Groupby.R
        icon: file-code
code-overflow: wrap
code-fold: show
editor: visual
df-print: kable
R.options:
  warn: false
code-tools: true
execute:
  eval: true
---
# Load the data

```{r} 
np_data <- read.csv("https://raw.githubusercontent.com/melaniewalsh/Neat-Datasets/main/1979-2020-National-Park-Visits-By-State.csv", stringsAsFactors = FALSE)
```

# Load dplyr library
```{r warning="ignore"} 
library("dplyr")
```
# Exercise 1

What is the average number of visits for *each state*?

Save as `avg_state_visits` and then view the resulting dataframe.

```{r} 
# Your code here
```

Discuss/consider: What state has the most and least average visits? What patterns or surprises do you notice?

# Exercise 2
What is the average number of visits for *each National Park*?

Save as `avg_park_visits` and then view the resulting dataframe.

```{r} 
# Your code here
```

Discuss/consider: Which National Park has the most and least average visits? What patterns or surprises do you notice?

# Exercise 3: 
How many National Parks are there in *each state*?

Save your answer as `distinct_parks`.

```{r} 
# Your code here
```

Discuss/consider: What state has the most and least average visits? What patterns or surprises do you notice?

