---
title: "Visualize Author Gender By Year (Solution)"
date: "2024-02-26"
categories: [ggplot2, dplyr, line-plot, bar-plot, groupby, R, solution, exercise]
toc: true
format:
   html: default
code-overflow: wrap
code-fold: show
editor: visual
df-print: kable
R.options:
  warn: false
code-tools: true
execute:
  eval: true
  message: false
  warning: false
---

These exercises explore data about African American periodical poetry from 1900–1928, originally compiled by Amardeep Singh and Kate Hennessey at Lehigh University. For more context about the dataset, see the [data essay](../aa-periodical-poetry.qmd).

**Concepts covered:**

- Groupby with multiple columns
- Value counts within groups
- Line plots and bar plots with categorical coloring (ggplot2)

# Load libraries

```{r}
library(dplyr)
library(ggplot2)
```

# Read in CSV

```{r}
aa_df <- read.csv("https://raw.githubusercontent.com/melaniewalsh/responsible-datasets-in-context/main/datasets/aa-periodical-poetry/African-American-Periodical-Poetry_1900-1928-Created-by-Amardeep-Singh-and-Kate-Hennessey,-Lehigh-University.csv",
  stringsAsFactors = FALSE)
```

# View first 5 rows

```{r}
head(aa_df)
table(aa_df$gender..if.known.)
```

# Group by year, count instances of author by gender

```{r}
aa_gender_by_year <- aa_df %>%
  group_by(year, gender..if.known.) %>%
  summarize(count = n(), .groups = "drop")

aa_gender_by_year
```

# Visualize with ggplot2 - Line Plot

```{r}
ggplot(aa_gender_by_year, aes(x = year, y = count, color = gender..if.known.)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1900, 1930, by = 5)) +
  labs(title = "Author Gender by Year",
       x = "Year",
       y = "Count",
       color = "Gender (if known)") +
  theme_minimal()
```

# Visualize with ggplot2 - Bar Plot

```{r}
ggplot(aa_gender_by_year, aes(x = factor(year), y = count, fill = gender..if.known.)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Author Gender by Year",
       x = "Year",
       y = "Count",
       fill = "Gender (if known)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
