---
title: "Norton Anthology & SPL Checkouts Data Exploration (Solution)"
date: "2026-02-25"
categories: [dplyr, exercise, solution]
format:
  html: default
code-overflow: wrap
code-fold: show
editor: visual
df-print: kable
R.options:
  warn: false
code-tools: true
execute:
  eval: true
---

# <span style="color:red;"> Solution </span>

## Norton Anthology & SPL Checkouts Data Exploration

<span style="color:green;"> [Exercise Without Solutions](Norton-Anthology-SPL-Checkouts-Exercise.qmd) </span>

These exercises explore checkout data from the Seattle Public Library for authors in the post-1945 volumes of the *Norton Anthology of American Literature*, one of the most widely used literary anthologies in U.S. college classrooms. The dataset was compiled as part of the research paper ["The Canon in Circulation"](https://doi.org/10.63744/P6qPH135jhY2) (Gupta et al., 2025). The code and data are available on [GitHub](https://github.com/melaniewalsh/Canon-in-Circulation), and you can explore more of the analysis [here](https://melaniewalsh.github.io/whats-seattle-reading/posts/norton-anthology-american/). For more context, see the [Library Checkouts for the Top 500 "Greatest" Novels data essay](../index.qmd).

**Concepts covered:**

- Groupby and aggregation (sum of checkouts)
- Sorting and ranking (top N values)
- Horizontal bar charts
- Filtering by a specific author
- Time series line plots (monthly checkouts over time)

# Load the data

```{r}
#| message: false
library(dplyr)
library(ggplot2)

norton_df <- read.csv("https://seattle-library-checkout-data.s3.us-west-2.amazonaws.com/norton-anthology_spl-checkouts_2005-2025.csv",
  stringsAsFactors = FALSE)
head(norton_df)
```

# Exercise 1

Find the top 10 authors by total checkouts in the Norton Anthology SPL checkouts dataset. Create a horizontal bar chart to visualize the results.

Save the top authors as `top_authors`.

```{r}
top_authors <- norton_df %>%
  group_by(Normalized.Creator) %>%
  summarize(total_checkouts = sum(Checkouts)) %>%
  arrange(desc(total_checkouts)) %>%
  head(10)

ggplot(top_authors, aes(x = reorder(Normalized.Creator, total_checkouts), y = total_checkouts)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 10 Norton Anthology Authors by SPL Checkouts",
       x = "Author", y = "Total Checkouts") +
  theme_minimal()
```

Discuss/consider: Which authors from the Norton Anthology are most popular at the Seattle Public Library? Are these authors you expected to see at the top?

# Exercise 2

Find the top 10 titles by total checkouts. Create a horizontal bar chart to visualize the results.

Save the top titles as `top_titles`.

```{r}
top_titles <- norton_df %>%
  group_by(Normalized.Title) %>%
  summarize(total_checkouts = sum(Checkouts)) %>%
  arrange(desc(total_checkouts)) %>%
  head(10)

ggplot(top_titles, aes(x = reorder(Normalized.Title, total_checkouts), y = total_checkouts)) +
  geom_bar(stat = "identity", fill = "darkorange") +
  coord_flip() +
  labs(title = "Top 10 Norton Anthology Titles by SPL Checkouts",
       x = "Title", y = "Total Checkouts") +
  theme_minimal()
```

Discuss/consider: Which titles are checked out most frequently? How do these compare to the most popular authors?

# Exercise 3

Create a time series line plot showing monthly checkouts for Octavia E. Butler over time.

Filter the data for Octavia E. Butler, group by year and month, and plot the results.

```{r}
butler_df <- norton_df %>%
  filter(Normalized.Creator == "Octavia E. Butler") %>%
  group_by(CheckoutYear, CheckoutMonth) %>%
  summarize(monthly_checkouts = sum(Checkouts), .groups = "drop") %>%
  mutate(date = as.Date(paste(CheckoutYear, sprintf("%02d", CheckoutMonth), "01", sep = "-")))

ggplot(butler_df, aes(x = date, y = monthly_checkouts)) +
  geom_line(color = "purple") +
  labs(title = "Monthly SPL Checkouts for Octavia E. Butler",
       x = "Date", y = "Checkouts") +
  theme_minimal()
```

Discuss/consider: What patterns do you notice in the checkout trends for Octavia E. Butler? Are there any notable spikes or dips? What might explain them?
