---
title: "SPL Top 500 Data Exploration (Solution)"
date: "2026-02-25"
categories: [dplyr, exercise, solution]
format:
  html: default
code-overflow: wrap
code-fold: show
editor: visual
df-print: kable
R.options:
  warn: false
code-tools: true
execute:
  eval: true
---

# <span style="color:red;"> Solution </span>

## SPL Top 500 Data Exploration

<span style="color:green;"> [Exercise Without Solutions](SPL-Top-500-Data-Exploration-Exercise.qmd) </span>

These exercises explore checkout data from the Seattle Public Library for the Top 500 "Greatest" Novels — the novels most widely held in libraries according to OCLC. For more context about the dataset, see the [data essay](../index.qmd).

**Concepts covered:**

- Groupby and aggregation (sum of checkouts)
- Sorting and ranking (top N values)
- Time series line plots (monthly checkouts over time)
- String filtering (finding titles that contain a keyword)
- Pivot tables
- Correlation matrices and heatmaps — a correlation coefficient measures how closely two variables move together, ranging from -1 (perfect inverse relationship) to 1 (perfect positive relationship), with 0 meaning no relationship

# Load the data

```{r}
#| message: false
library(dplyr)
library(ggplot2)
library(tidyr)

spl_df <- read.csv("https://responsible-datasets-in-context.s3.us-west-2.amazonaws.com/top_500_spl_df.csv",
  stringsAsFactors = FALSE)
head(spl_df)
```

# Exercise 1

Find the top 10 authors and top 10 books by total checkouts in the SPL Top 500 dataset. Display them as tables.

Save the results as `top_authors` and `top_books`.

```{r}
top_authors <- spl_df %>%
  group_by(author) %>%
  summarize(total_checkouts = sum(checkouts)) %>%
  arrange(desc(total_checkouts)) %>%
  head(10)

top_authors
```

```{r}
top_books <- spl_df %>%
  group_by(top_500_title) %>%
  summarize(total_checkouts = sum(checkouts)) %>%
  arrange(desc(total_checkouts)) %>%
  head(10)

top_books
```

Discuss/consider: Which authors and books are most popular at the Seattle Public Library? Are there any surprises?

# Exercise 2

Create a time series line plot of monthly checkouts for "Pride and Prejudice" over time.

Filter the data for "Pride and Prejudice", group by year and month, and plot the results.

```{r}
pride_df <- spl_df %>%
  filter(top_500_title == "Pride and Prejudice") %>%
  group_by(checkoutyear, checkoutmonth) %>%
  summarize(monthly_checkouts = sum(checkouts), .groups = "drop") %>%
  mutate(date = as.Date(paste(checkoutyear, sprintf("%02d", checkoutmonth), "01", sep = "-")))

ggplot(pride_df, aes(x = date, y = monthly_checkouts)) +
  geom_line(color = "darkgreen") +
  labs(title = "Monthly SPL Checkouts for Pride and Prejudice",
       x = "Date", y = "Checkouts") +
  theme_minimal()
```

Discuss/consider: What patterns do you notice in the checkout trends? Are there any seasonal patterns or notable changes over time?

# Exercise 3

Calculate the correlation between the monthly checkout patterns of Harry Potter books and display the results as a heatmap.

Filter for Harry Potter titles, pivot the data so each book is a column, compute the correlation matrix, and visualize it.

```{r}
hp_df <- spl_df %>%
  filter(grepl("Harry Potter", top_500_title, ignore.case = TRUE)) %>%
  group_by(top_500_title, checkoutyear, checkoutmonth) %>%
  summarize(total_checkouts = sum(checkouts), .groups = "drop") %>%
  mutate(date = paste(checkoutyear, sprintf("%02d", checkoutmonth), sep = "-")) %>%
  select(top_500_title, date, total_checkouts) %>%
  pivot_wider(names_from = top_500_title, values_from = total_checkouts, values_fill = 0)

cor_matrix <- cor(hp_df %>% select(-date), use = "complete.obs")

cor_long <- as.data.frame(as.table(cor_matrix))
names(cor_long) <- c("Book1", "Book2", "Correlation")

ggplot(cor_long, aes(x = Book1, y = Book2, fill = Correlation)) +
  geom_tile() +
  geom_text(aes(label = round(Correlation, 2)), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Correlation Between Harry Potter Book Checkouts",
       x = "", y = "")
```

Discuss/consider: Which Harry Potter books have the most correlated checkout patterns? What might explain these correlations?
